### 2018-05-07 doc/incertae_sedis.tex
up to naming section

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\linespread{1.5}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newenvironment{lyxcode}
    {\par\begin{list}{}{
        \setlength{\rightmargin}{\leftmargin}
        \setlength{\listparindent}{0pt}% needed for AMS classes
        \raggedright
        \setlength{\itemsep}{0pt}
        \setlength{\parsep}{0pt}
        \normalfont\ttfamily}%
     \item[]}
    {\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\sloppy
%\setlength{\parskip}{6pt}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}
\makeatother

\usepackage{babel}
\usepackage{xspace}
\usepackage{color}
\usepackage[colorlinks,linkcolor=blue,citecolor=blue]{hyperref}
\begin{document}

% \raggedright

\global\long\def\taxonomy{\mbox{\ensuremath{\mathbb{T}}}}

\global\long\def\prunedTaxonomy{\taxonomy_{P}}

\global\long\def\phyloinputs{\mathcal{T}}

\global\long\def\expandedPhylo{\phyloinputs_{E}}

\global\long\def\summaryTree{\mathbb{S}}

\global\long\def\prunedSummary{\summaryTree_{P}}

\global\long\def\collections{\mathcal{C}}

\newcommand{\note}[1]{}

\newcommand{\mthcomment}[1]{{\color{red} #1}\xspace}

\newcommand{\bdrcomment}[1]{{\color{cyan} #1}\xspace}

\title{\texttt{Taxonomic supertree construction with}\texttt{\emph{
Incertae sedis}}\texttt{ taxa}}

\author{Benjamin D.~Redelings and Mark T.~Holder} 
\maketitle

\section{Introduction}

Supertree methods combine a collection of input trees with different taxon sets
into a single tree on the combined taxon set. These methods are usually applied
to leaf-labeled phylogenetic estimates.  Here we describe a supertree approach
that overcomes issues associated with using a taxonomy as one of the inputs to
a supertree method.

The Open Tree of Life project \citep{HinchliffEtAl2015} seeks to build a
comprehensive supertree \citep[see][]{redelings2017supertree} covering all of
life. The approach is to combine information in published phylogenies with a
comprehensive taxonomy that supplies taxon names. The supertree method has
been used with the Open Tree Taxonomy \citep[OTT hereafter, see][]{rees2017automated},
but could be used in more general contexts. The taxonomy is a crucial input for
several reasons: it covers a wide range of species, its list of names allows
for alignment of different phylogenetic estimates to each other, and it provides
names for clades. Only a small proportion of known species have been included in
a phylogenetic analysis, thus a taxonomy is important for achieving
comprehensive coverage of known taxa. Phylogenetic estimates collected from the
published literature often use different names for the same species. Lists of
synonyms in OTT (and other taxonomies) allow data curators to ``align'' input
phylogenies to the taxonomy \citep[see][ for discussion of the curation process
that the Open Tree of Life project uses to align phylogenetic estimates to OTT]{McTavishEtAt2015}.
This alignment is important for recognizing when two different estimates refer
to the same taxon. Biologists are familiar with names for ``higher'' taxa (taxa
above the species rank). Thus, naming clades in the supertree makes the tree
easier to navigate and use.

Some taxonomies indicate which taxa are thought to be the result of
hybridization, but the supertree approach described here does not consider any
hybridization. Thus the taxonomic hierarchy can be converted to a tree. While
not all taxonomists believe that named taxa should correspond to clades \citep[see,
for example,][]{HorandlS2010}, the priniciples of phylogenetic classification
are used by such a large majority of practicing taxonomists, that we have chosen
to treat the taxonomic tree that mirrors the hierarchy of the taxonomy as an
estimate of phylogenetic relationships. We refer to labels of taxonomy nodes as
``taxon names'', and assume that taxon names and taxonomy nodes have a
one-to-one correspondence. Note that the taxonomy may also contain out-degree
1 nodes. These nodes correspond to monotypic taxa which contain a single child
of lower taxonomic rank.

The supertree method described by \citet{redelings2017supertree} is able to use
a set of phylogenetic estimates and taxonomy to produce a supertree with clades
named according to the taxonomy. However, the method was not able to accommodate
the fact that taxonomies frequently contain nodes with uncertain placement.
These nodes are often labelled ``\emph{incertae sedis}'', which means
``uncertain seat'' in Latin. The common interpretation of such taxa is that
they may not be moved further towards the root, but may be moved into their
sibling taxa. For example, a genus with a sibling that is a family may be
annotated as ``\emph{incertae familia}'', indicating that it is unclear which
family the genus should be placed in. \emph{Incertae sedis} taxa frequently
occur when a specimen is identified down to a given taxonomic level, but no
further. Extinct taxa are often \emph{incertae sedis}. OTT is constructed from
several source taxonomies, and these sources include various ways of indicating
that the position of a taxon within the hierarchy is uncertain. OTT uses a
series of ``flags'' to annotate these taxa. The supertree of
\citet{redelings2017supertree} simply pruned these taxa from the taxonomy and
input trees. Thus, the final supertree lacked taxa which were \emph{incertae
sedis} or the descendants of such taxa.

Here we describe a supertree method so that we can resolve taxonomic uncertainty
by using published phylogenies to place \emph{incertae sedis} taxa. Developing
this method required deciding on a set of operational semantics to be used when
interpreting the \emph{incertae sedis} annotation. These semantics affect the
interpretation of the phylogenetic statements being made by the input taxonomy
and the rules for applying taxonomic names to the final supertree. In addition
to discussing how to interpret the \emph{incertae sedis} annotation, we describe
the algorithmic changes to the supertree pipeline that were required to
adequately represent the taxonomic uncertainty. The approach described here
allows us to include thousands of new taxa in our supertree analysis that were
previously filtered out. It will also enable the Open Tree of Life project to
include extinct taxa, since many of these taxa are \emph{incertae sedis}. This
makes substantial progress towards our goal of comprehensive inclusion of known
taxa into the Open Tree of Life summary tree.

\subsection{Background}
\label{sec:background}
\paragraph{Previous algorithm}

The supertree algorithm of \citet{redelings2017supertree} takes as input a
ranked list $\phyloinputs=\left\{ T_{1},\ldots,T_{n}\right\} $ of leaf-labeled rooted input phylogenies and a
single rooted taxonomy tree $\taxonomy_{p}$ that is ranked below all of the input trees.
As discussed in \citet{redelings2017supertree}, the preferred tree would
display as many of the highly ranked input splits as possible. If computational
barriers were not an issue, the method would act by first producing a ranked
list of rooted splits: $\mathcal{S}=S(T_{1})+S(T_{2})+\ldots+S(T_{n})+S(\mathbb{T})$ where $S(T_{i})$ denotes a list
of non-trivial rooted splits created from a post-order traversal of tree $i$ and '$+$'
denotes concatenation. Then the supertree would be produced in a greedy manner
by iterating over the list $S$, and adding each rooted split to a growing rooted
supertree if that split is compatible with the supertree. In other words, the set
$C$ is initialized to the empty set, and we then iteratively consider each split in
the list $\mathcal{S}$ and insert it into $C$ if the resulting set remains jointly compatible.
The BUILD algorithm \citep{AhoSSU1981} is used both to check compatibility,  and
to construct many aspects of a compatible tree from the final set $C$.  Refer to
\citet{redelings2017supertree} for details of the procedure, including a
description of an optimization for pruning taxa that do not occur in any
phylogenetic estimates during pre-processing and then reattaching them to a
backbone tree.  This optimization is not crucial to the current work, but it
does cause some groups to attach more tipward than they would if the entire tree
were constructed using BUILD.

Because of the large number (over 2.6 million) of tips in the full tree, the
supertree method of \citet{redelings2017supertree} uses an approximation
approach that relies on decomposition of the full problem into subproblems.
The decomposition partitions the list of taxonomy splits $S(\taxonomy_{p})$ into two lists:
$\mathcal{Z}(\taxonomy_{p}),$ the list of splits from the taxonomy tree which are compatible with every
rooted split in the phylogenetic inputs, and $\mathcal{Y}(\taxonomy_{p})$, the list of taxonomic splits
that conflict with at least one phylogenetic input split. Conceptually the
approximate algorithm works by greedily adding all compatible splits in
altered ranked list of splits: $\mathcal{S}^{\prime}=\mathcal{Z}(\taxonomy_{p})+S(T_{1})+S(T_{2})+\ldots+S(T_{n})+\mathcal{Y}(\taxonomy_{p})$.

In practice, the decomposition allows the compatibility of splits to be checked
on smaller, relabeled trees. This can be done much more efficiently than if
compatibility was checked using the BUILD algorithm of \citet{AhoSSU1981} on the
full leaf set. Thus, the supertree algorithm first divides the supertree problem
into independent pieces by bisecting trees at the uncontested taxa which
correspond to $Z$, then solves the supertree problem on these pieces, and finally
glues the pieces back together to produce the full tree.

The taxonomy tree, $\taxonomy_{p}$, is derived from OTT. As mentioned above, OTT not only
contains a taxonomic hierarchy, but also contains taxonomic labels and
annotations (referred to as ``flags'') on the taxa to carry extra information.
The entire hierarchy encoded by OTT is a complete taxonomic tree, $\taxonomy^{\ast}.$ Nodes in
$\taxonomy^{\ast}$ have a one-to-one correspondence with a set of labels $\mathcal{N}$ that are called
taxon names. We will decompose $\mathcal{N}$ into the taxon names for leaf taxa $\mathcal{L}$ and
higher taxa $\mathcal{H}$. We will use $\mathcal{U}$ to refer to the subset ($\mathcal{U}\subseteq\mathcal{N}$) of taxa that are
annotated with the \emph{incertae sedis }property. Other annotations identify
dubious taxa and taxonomic names that are unwanted artifacts of the process of
building OTT. Thus, the taxonomic tree which is used as an input to the
supertree algorithm is pruned version of $\taxonomy^{\ast}$. For our purposes, the relevant
information from the taxonomy consists of the taxonomy tree $\taxonomy^{\ast}$, the labels $\mathcal{N}$,
and set $\mathcal{U}$ of \emph{incertae sedis} taxa. In previous work, taxa in $\mathcal{U}$ and all
of their descendants were all pruned from the taxonomic hierarchy when $\taxonomy_{p}$ is
extracted from OTT. In the present work, we consider supertree approaches that
operate on a taxonomic tree $\taxonomy$, which is produced from $\taxonomy^{\ast}$but does not prune
\emph{incertae sedis} taxa.

Throughout we assume that leaf labels on input phylogenetic trees are only taken
from the leaf set of the taxonomic tree that is being used and that each label
occurs at most once in each input tree. \citet{redelings2017supertree} describe
a method called ``exemplification'' that ensures this is the case. Thus pruning
of the \emph{incertae sedis} taxa from the taxonomy was also accompanied by
pruning those taxa and all of their descendants from the input phylogenetic
trees.

\begin{figure}
\subfloat[Input tree \#1]{\includegraphics[scale=0.5]{Figures/is-clade/tree1-shaded}

}\hfill{}\subfloat[Taxonomy]{\includegraphics[scale=0.5]{Figures/is-clade/tax-shaded}

}

\subfloat[Synthesis Tree (no \emph{incertae sedis})]{\includegraphics[scale=0.5]{Figures/is-clade/synth1-no-is-shaded}

}\hfill{}\subfloat[Synthesis Tree (\emph{incertae sedis} aware)]{\includegraphics[scale=0.5]{Figures/is-clade/synth1-shaded}

}

\caption{\label{fig:Handling-incertae-sedis}Handling \emph{incertae sedis
}taxa recovers additional edges and taxon names.
Supertree construction on input tree (a) and taxonomy tree (b) with \emph{incertae sedis}
clade $C$ leads to synthesis tree (c) if \emph{incertae sedis} annotations are
ignored, but to synthesis tree (d) if \emph{incertae sedis} annotations are respected.
The tree in (d) recovers taxon names $AB$ and $A$, as well as the edge to clade $AB$
that groups $A$ and $B$. Taxon names and edges that are conditional on handling
\emph{incertae sedis}
are shaded grey.
}

\end{figure}

\subsection{A naive semantics: consequences of ignoring the annotation}

The most straightforward approach for increasing the taxonomic completeness of
the supertree is simply to no longer prune taxa from $\taxonomy$ because they are flagged
as \emph{incertae sedis}, and then run the rest of the pipeline of
\citet{redelings2017supertree} unaltered.  Before we propose a semantics of
\emph{incertae sedis}, it is worth pointing out what this naive semantics is lacking.

This approach incorrectly treats the intrusion of an \emph{incertae sedis} taxon
into a sibling taxon as a conflicting with monophyly of the sibling taxon. For
example, in Fig. \ref{fig:Handling-incertae-sedis}, the placement of the
\emph{incertae sedis} taxon $C$ within $A$ and $AB$ is considered
to contradict the monophyly of $A$ and $AB$ and thus to conflict with those taxa.
This has three main effects.  These effects are illustrated in the difference
between the synthesis tree in Fig. \ref{fig:Handling-incertae-sedis}c, which
ignores \emph{incertae sedis} annotations, and the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}d, which respects them.

First, the naive approach results in a less-resolved tree.  While input trees are
free to place \emph{incertae sedis} taxa within a sibling, the naive approach
interprets this placement as conflict with the sibling.  Since the taxonomy
is ranked below all input trees, conflict is always resolved in favor of the
input trees, and so the sibling in the taxonomy is removed from the synthesis
tree.  For example, in the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}c, the placement of taxon $C$ in group $AB$ is
interpreted as a rejection of the monophyly of $AB$. The edge leading to $AB$ is
therefore not incorporated into the synthesis tree.  In contrast, the synthesis
tree in Fig. \ref{fig:Handling-incertae-sedis}d that does not reject the
monophyly of $AB$ retains the grouping of $A$ and $B$.

Second, the naive approach results in the removal of clade names for all clades that
have an \emph{incertae sedis} taxon placed within them.  For example, in Fig.
\ref{fig:Handling-incertae-sedis}c the names for the clades $A$ and $AB$ are lost.

Third, the naive approach cannot use information from phylogenies to resolve the
location of uncertain taxa.  For example, the synthesis tree in
\ref{fig:Handling-incertae-sedis}c has lost the label for $A$, and so does not
allow us to note that $C$ has been placed within $A$.  In contrast, the synthesis
tree in Fig. \ref{fig:Handling-incertae-sedis}d retains the label for $A$, which
allows us to resolve taxonomic uncertainty by inferring that $C$ has been placed
within $A$.

\section{\label{sec:Semantics-of-incertae}Semantics of \emph{incertae
sedis} taxa}

\subsection{Goals for an \emph{incertae sedis} semantics}

In order to incorporate \emph{incertae sedis }taxa into a supertree
analyses, we must arrive at an operational definition of the meaning
of the \emph{incertae sedis} label.
The core meaning of the phrase is
that the annotated taxon may actually be a member of one of the taxa
that appear as siblings to it in the heirarchical representation of
the taxonomy.
In the most expansive interpretation, the author of the
taxonomy is stating that the \emph{incertae sedis} taxon could be
placed anywhere inside one of the sibling taxa or their descendants.
However, frequently only a subset of the possible placement points
would be viewed as plausible.

We seek a semantics for supertrees with \emph{incertae sedis }taxa
that satisfy the following properties:
\begin{enumerate}
    \item An \emph{incertae sedis} node may intrude into its siblings and their descendants without conflict.
    \item An \emph{incertae sedis} node has an unrestricted range within the descendants of its sibling taxa.
    \item The semantics of an \emph{incertae sedis} node does not depend on additional information such as taxon ranks, but only on the taxonomy tree.
    \item The semantics is based on deriving a rooted (partial) split for each branch of the taxonomy.
      % These could be goals, but maybe discuss them as side-effects:
      % Side effect #1. Don't interdigitate
      % Side effect #2. Allow placing A next to B, and then placing B within A. (See Fig. 6)
\end{enumerate}

\subsection{Terminology for rooted splits}

The rooted split associated with taxonomy node $n$ serves to determine both when $n$ is in conflict
with an input tree, and to where to place the taxon name for clade $n$ on the synthesis tree.
% Removed mention of incertae sedis possibly affecting subproblem decomposition.  It
% actually doesn't but that is surprising, and therefore kind of a result.
% \paragraph{Definition of rooted splits}
As is typically done, we define rooted splits for each node by noting that
removing the edge connecting the node to its parent would produce two trees.
The leaf label sets of these trees
the partition the full leaf set $\mathcal{L}$ into two groups.
The include set $\mathcal{I}(n)$ which
does not contain the root, and the exclude set $\mathcal{E}(n)$ which does contain the root.
Such a split may be written
\[ \mathcal{I}(n)|\bullet\mathcal{E}(n).\]
where the $\bullet$ indicates the root. 
The include set is the set of labels for leaves that are
descendants of the node.
The exclude set is the set of
labels for leaves that are not descendants of that node. 
If no taxa are \emph{incertae sedis},
then the exclude set for a node is just the total tip set minus the include set
for the node: $\mathcal{E}(n)  =\mathcal{L}-\mathcal{I}(n)$.
%For a node $n$ on the tip-ward side of an edge $e$, we
%may also write $\mathcal{I}(n)$ for $\mathcal{I}(e)$, and $\mathcal{E}(n)$ for $\mathcal{E}(e)$.
For any two rooted splits $A=A_{1}|\bullet A_{2}$
and $B=B_{1}|\bullet B_{2}$ , we say that $A$ \emph{displays }$B$ if $B_{1}\subseteq A_{1}$ and $B_{2}\subseteq A_{2}$.

Note that when a taxon rooted at node $n$ is not marked as \emph{incertae sedis},
then the leaf labels that are descendants of $n$ will be in the include set for
$n$ and all its ancestors and will be in the exclude set for all other nodes.

\paragraph{Equivalence of a taxonomy tree and a list of rooted splits}

Note that, for the purposes of the supertree method, all of the information from
the taxonomic tree is contained in the set of rooted splits. Thus, one could
imagine replacing that taxonomy, $\taxonomy_{p}$, with a set of rooted splits $\mathcal{R}(\taxonomy_{p})$ such that
each edge in $\taxonomy_{p}$ is encoded in one split in $\mathcal{R}(\taxonomy_{p})$. If the order of trees is set
to be: $\mathcal{R}(\taxonomy_{p})=\mathcal{Z}(\taxonomy_{p}) + 
\mathcal{Y}(\taxonomy_{p})$, then using $\mathcal{R}(\taxonomy_{p})$ in place of $\taxonomy_{p}$ in the supertree
algorithm would produce the same output.

\paragraph{Equivalence rooted split definitions and phylogenetic definitions of names}
In terms of phylogenetic nomenclature the supertree pipeline acts as if the
taxonomy defines a conditional node-based name \citep[see][]{deQueiroz2013} for each
higher taxon.
So for the name labeling node $n$ in the taxonomic
tree, the operational definition would be: ``the clade defined by the most recent
common ancestor (MRCA) of $\mathcal{I}(n)$ as long as that MRCA is not an ancestor of
$\mathcal{E}(n)$.''
OTT is not based on a set of explicitly defined phylogenetic name definitions.
In many contexts, a branch-based definition for $n$ (``the clade by all of the taxa more
closely related to all fo them members of $\mathcal{I}(n)$ than to any member of
$\mathcal{E}(n)$, if such a clade exists'') would result in similar behavior
during the tree construction. 
An exception to this (discussed below) arises in the step of assigning names to the 
    supertree.

\subsection{Split-based semantics for \emph{incertae sedis} taxa}

An \emph{incertae sedis} taxon $n$ differs from a normal taxon because it can be
placed as a descendant of some set of taxa without invalidating these taxa.
We will use $\mathcal{P}_n$ to denote this set of taxa which are {\emph{potential}} 
ancestors for \emph{incertae sedis} taxon $n$.
Note that this distinct from the taxa that are ancestral to $n$ in the taxonomy;
    those are taxa which must contain all of the descandants of $n$.
The effect of the \emph{incertae sedis} annotation of taxon $n$ is
    to simply remove $\mathcal{I}(n)$ from the exclude set for each $p\in\mathcal{P}_n$.
The include sets of these splits remain unchanged, but the exclude sets may be reduced
\emph{incertae sedis} taxa to intrude.
Such a split does
not mention the complete set of tip labels $\mathcal{L}$ and is therefore called a
`partial split'', as opposed to a ``full split'' that contains the entirely of $\mathcal{L}$.

When considering \emph{incertae sedis} annotations, the taxonomy is still comprehensive 
if we include the degenerate grouping for the base
    node $\rho$ which has $\mathcal{I}(\rho) = \mathcal{L}$ and $\mathcal{E}(\rho) = \emptyset$,
    but many of the taxonomic nodes will correspond to partial splits.
Thus, when interpreted correctly the \emph{incertae sedis} 
annotations are reducing the amount of information that the taxonomy brings to
the supertree problem.

%  of any of its siblings.
% Thus, annotating $n$ as
% \emph{incertae sedis} really modifies the meanings of the taxanomic statements (rooted splits)
% of other
% that of the siblings of $n$ and all taxonomic nodes that descend from those siblings.
%  so that they
% no longer exclude $n$.
% %In terms of rooted splits, we can represent this lack of
% %   exclusion by omitting taxa from the exclude set for a node.
% % We seek to define modified splits for each branch of the taxonomy tree.
%  into their sibling taxa.
% % Therefore, a semantics can be given simply in terms of the exclude set for each taxonomy node.

\subsection{Unrestricted range and ignoring additional information}
In principle, an \emph{incertae sedis} taxon $n$ could carry with it additional
    information that indicates which specific points within the taxonomic tree
    it could attach.
Unfortunately, reliable information about the range of valid attachment points
for each \emph{incertae sedis} taxon is lacking in OTT and most other large-scale
taxonomies.
Constructing a range for each incertae sedis taxon relies on detailed
readings of character argumentation (which is lacking in OTT) or rank-based
information (which is frequently untrustworthy in OTT). Thus, here we pursue
an approach of ascribing a meaning to the \emph{incertae sedis} label which
attempts to capture the core of the idea that it articulates, but which can be
applied automatically across the taxonomy without additional information.

Specifically, the methods described here only consider cases in which 
    $\mathcal{P}_n$ can be determined from the structure of the tree.
Our method takes $\mathcal{P}_n$ to be the set of nodes that
    are in the taxonomic subtrees rooted at the sibling taxa of \emph{incertae sedis} taxon $n$.
This allows us to encode the exclude set of some node $x$ efficiently by sweeping
    over the tree from root to tip.
If $\mathcal{S}(x)$ is the set of siblings of $x$ which are not annotated as being 
    \emph{incertae sedis} and $p(x)$ is the parent of $x$, then
\begin{align}
    \mathcal{E}(x) = \mathcal{E}(p(x)) \cup \bigcup_{s\in \mathcal{S}(x)} \mathcal{E}(s)
\end{align}
with $\mathcal{E}(\rho) = \emptyset$ for the root of the taxonomic tree.

%If $\mathcal{P}$ is the set of taxa that are potential ancestors for \emph{incertae
%  sedis} taxon $x$, then we may simply remove the descendants of $x$ from the exclude
%  set for each $p\in\mathcal{P}$.  Such an approach would require that the path from $x$ to $p$
%  go through only other potential ancestors.

% Supporting more precise information about $\mathcal{P}_n$  for each \emph{incertae sedis}
%    taxon $n$ is an important future direction to be explored.
% For example, a genus
% might be characterized as \emph{incertae sedis} at the level of the subfamily.
% Thus, it could be placed inside any of the hypothesized sub-families, or within
% its own subfamily, but it would be unexpected for it to be placed inside one of
% the other genera.

% In order to determine how the exclude sets should be modified to account for
% \emph{incertae sedis} taxa, we first attempt to express the unmodified exclude
% sets in a helpful way.  We note that a node $n$ ordinarily excludes the
% descendants of $\mathcal{I}(s)$ for all of its siblings, as well as the siblings
% of all of its ancestors.  This leads to the following equation:
% \begin{align}
% \mathcal{E}(n) & = \mathcal{L} - \mathcal{I}(n)\nonumber\\
%  & =\left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a)\right]\nonumber\\
%  & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a)\right]\nonumber\\
%  & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\mathcal{E}(parent(n)).\label{eq:exsetformonea-traditional}
% \end{align}

% When modifying the exclude set for a node $n$, we seek to find all the \emph{incertae
% sedis} taxa that may intrude into it. It turns out that the incertae sedis taxa that
% might intrude are just the \emph{incertae sedis} siblings of $n$ and all of its ancestors.
% \begin{align}
% \mathcal{E}(n) & =
%     \left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\nonumber\\
% & =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\nonumber\\
% & \,\,\,\,\,\,\,\,\,\,\,\,\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\nonumber\\
% & =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\mathcal{E}(parent(n)).\label{eq:exsetformonea}
% \end{align}
% This formula allows us to compute exclude sets via a pre-order traversal on the
% taxonomy tree with the exclude set for the root node defined to be $\emptyset$.

\subsection{Naming }

\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/solution-shaded}}
\hfill{}
\caption{\label{fig:One-name-can}One name can be consistent with multiple nodes.}
\end{figure}

After constructing a supertree, tip nodes already have names in \emph{$\mathcal{L}$}.
However, we still need to assign higher taxon names to internal supertree nodes
based on the taxonomy tree in the problem. Each taxon name $n$ corresponds to a
split $S(n)=S(n)_{1}|\bullet S(n)_{2}$ on the corresponding branch of the taxonomy tree. Without
\emph{incertae sedis}, such splits are always of the form $S(n)_{1}|\bullet\mathcal{L}-S(n)_{1}$, but
with \emph{incertae sedis} taxa $S_{2}(n)$ may be smaller than $\mathcal{L}-S(n)_{1}$.

Without \emph{incertae sedis}, each name applies to at most one node, and each
node can take at most one name, with the exception of monotypic taxa. Thus, we
may simply search the solution tree for a node that has the same cluster $S(n)_{1}$
and apply the name $n$ to that node.

However, in the \emph{incertae sedis} framework, it is possible for one name to
apply to multiple nodes. For example, in Figure \ref{fig:One-name-can}b, the
name $A$ can apply to nodes $x$ and $y$. Here the name $A$ corresponds to the split
$A1\,A2\,|\bullet B\,C$, leaving out $D$ since it is \emph{incertae sedis}. The two nodes $x$
and $y$ display the splits $A1\,A2\,|\bullet B\,C\,D$ and $A1\,A2\,D|\bullet
B\,C$ respectively, and both
of these splits display the split $A1\,A2\,|\bullet B\,C$, so the name $A$ can apply to both
$x$ and $y$. This cannot happen without\emph{ incertae sedis} taxa except at
monotypic nodes. When faced with a choice about where to place a name, our
solution is to find the most tip-ward node where the name can apply and attach
the name to this node. This corresponds to a conservative choice about whether
\emph{incertae sedis} taxa should be added to an existing taxon.

\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/synth-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/solution2-shaded}}
\hfill{}
\caption{\label{fig:Multiple-names-can}Multiple names can apply to a single node.}
\end{figure}

It is also possible for multiple names to apply to a single node. For example,
in Figure \ref{fig:Multiple-names-can}a, the taxon $A$ corresponds to the split
$B1\,B2\,C\,|\bullet\,Y$, and its child taxon $B$ corresponds to the split $B1\,B2\,|\bullet\,Y$. The
edge leading to $A$ is consistent with the split for $B$, but the name $B$ is
applied to the node with the smallest include group. However, in the
solution tree (Fig. \ref{fig:Multiple-names-can}b), there is is only node for
both names $A$ and $B$ to apply to.  This kind of situation arises when a taxon
becomes monotypic after its \emph{incertae sedis} children are placed elsewhere.
For such cases, we can solve this problem by (i) selecting a taxon at the
node that is ancestral to all other taxa and (ii) moving this taxon to a
newly-introduced out-degree-1 parent.  For example, $A$ is moved to an
out-degree-1 parent in Fig. \ref{fig:Multiple-names-can}c.  We repeatedly perform
this procedure until no taxon at the node is the parent of all taxa.

However, there are cases where thus procedure does not result in one name per
node.  This case can occur, for example, when two incertae sedis siblings are 
``inter-digitated'', or  combined into a single clade in such a way that their
members cannot be separated.  When the members of the two clades share the same
MRCA, as in Fig. \ref{fig:interdigitation}, then the names for both clades
apply to the same node.  However, in this case, neither name is ancestral to
the other, and the previous solution does not apply.  In such cases, we choose
one of the names arbitrarily.

\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/is-interdigitation/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/is-interdigitation/synth-shaded}}
\hfill{}
\caption{\label{fig:interdigitation}Interdigitation of \emph{incertae sedis} siblings.  Here the names $A$ and $B$ apply to the same node on the solution tree, but neither is ancestral to the other on the taxonomy.}
\end{figure}

\subsection{Placement}\label{subsec:Placement}

After we have attached taxon names to the synthesis tree, we would
like to interpret the position of these names in terms of placing
\emph{incertae sedis }taxa in a revised taxonomy.
This would allow us
to interpret the synthesis tree as saying that phylogenetic
information has (for example) placed genus A within family B, or
perhaps outside of all named families.
The simplest approach to
placement involves noting whenever a taxon $B$ is a descendant of a
taxon $A$ on the named synthesis tree but not the taxonomy tree.
For
example, in Figure \ref{fig:Handling-incertae-sedis}, taxon $C$ is a
descendant of taxon $A$ in the synthetic tree 
(Fig.\ref{fig:Handling-incertae-sedis}d), but not the taxonomy 
(Fig.\ref{fig:Handling-incertae-sedis}b).
Thus, we could say that the synthetic tree places
$C$ within $A$. Furthermore, the synthetic tree additionally places
$C$ within $AB$.

A slightly more general case is shown in 
Figure \ref{fig:Broken-incertae-sedis}.
Here taxon $C1$ is an descendant of $A$ in the
synthesis tree, but not the taxonomy.
Thus $C1$ is placed within $A$.
We note that the most recent common ancestor (MRCA) of $A$ and $C1$ in
the taxonomy is the root node, and the path to the MRCA from $A$ is
$A\to AB\to root$, while the path to the MRCA from $C1$ is $C1\to C\to
root$.
Here the additional node $AB$ on the path to the MRCA from $A$
indicates that $C1$ is placed within $AB$ as well as $A$.
In contrast,
the additional node $C$ on the path to the MRCA indicates that $C1$ is
a broken \emph{incertae sedis} taxon.
Thus we may consider an
algorithm that, for each taxon $B$ on the synthesis tree, finds the
closest ancestral taxon $A$, and checks if $A$ is an ancestor of $B$
on the taxonomy.
If not, we compute the paths $A\to A_{1}\to\ldots\to
MRCA(A,B)$ and $B\to B_{1}\to\ldots\to MRCA(A,B)$.

However, this assumes that \emph{incertae sedis} taxa are always
placed within their siblings, or descendants of their siblings.
As
such, it does not handle cases such as Figure \ref{fig:Placing-one-incertae}, 
in which $I$ is first placed as sibling to $AI$, and then
$AI$ is placed within $I$.
Note that $I$ is not a sibling of $AI$ or a
descendant of a sibling.
Furthermore, the MRCA of $I$ and $AI$ is the
root, and the path from $AI$ to the root includes $A$, but $A$ is not
broken.
We suggest that this conundrum may be resolved by first
processing the placement of $I$ within $A$ by modifying the taxonomy
to reflect this placement\footnote{Does this actually work? Some
placements break taxa.}.
When we consider the placement of $AI$ within
$I$ on this modified taxonomy, the MRCA of $I$ and $AI$ is now $A$
instead of the root, and the correct interpretation results.
We
suggest that by walking the tree in a pre-order fashion and processing
shallowing placements first, the previous approach of finding the MRCA
and labelling intermediate nodes as either additional placements or
broken incertae sedis (or incertae\_sedis\_inherited) taxa holds.
Such
an approach should also handle cases containing nested incertae sedis
taxa, as in Figure \ref{fig:Placing-one-incertae}.

\begin{comment}
One question is whether this approach is able to handle nested
incertae sedis taxa.

Placement of incertae sedis taxa by input trees is unfortunately not
quite as simple as finding a single location where an I.S.
taxon
should attach.
For example, when an incertae sedis taxon is broken,
its children need to be ``placed'' separately.

Each input tree can relate to an \emph{incertae sedis} taxon
$(A,B,C)D$ in a number of ways
\begin{itemize}
    \item it could resolve
$A$, $B$, $C$, or $D$.
    \item it could place $D$ on a degree-2 (=out-
degree-1) node that bisects a branch 
    \item it could place a descendant
taxon of $D$ 
    \item it could place a descendant taxon of $D$ in a
\emph{different place} than another input tree.
    \item it could place
children of $D$ in multiple places, thus conflicting with the branch.
If the \emph{incertae sedis} taxon $(A,B,C)$ is broken, then $A$, $B$
and $C$ become \emph{incertae sedis} clades in their own right, that
may attach separately, except that they .
This is because none of $A$,
$B$, or $C$ is in the exclude set of the siblings of $D$.
\end{itemize}
\end{comment}

\section{The rooted-split semantics in practice}

In order to examine the results of the rooted splits-semantics above, we explore
the outcome on a series of complex cases.  These cases are intended to
illustrate some of the issues that arise when developing an operational semantics of
\emph{incertae sedis}, as well as the outcomes for the rooted-split semantics.
We think that they constitute one of the main contributions of the paper.  In
each case, the synthesis trees in this section are constructed using the
BUILD-based algorithm described in section \ref{sec:background}, without
decomposing them into smaller subproblems or pruning and then unpruning
taxonomy-only taxa.  Synthesis trees that are described as \emph{incertae
sedis}-aware apply the BUILD algorithm to splits constructed using equation
\ref{eq:exsetformonea}, whereas the synthesis trees that are not \emph{incertae
sedis}-aware are constructed from splits using equation
\ref{eq:exsetformonea-traditional}.

Much of the complexity that we explore here comes from the fact that it is not
only tip taxa that can be \emph{incertae sedis}.  It is also possible to have
higher ranked  \emph{incertae sedis} taxa, such as genera that are marked
\emph{incertae familia}.

\subsection{The \emph{incertae sedis} semantics restores taxonomic groups and names}

With the naive semantics, intrusion of an incertae sedis taxon into a sibling is
interpreted as conflict with that sibling.  This results in the loss of the edge
corresponding to the sibling taxon as well as the taxon name.

We see from Fig. \ref{fig:Handling-incertae-sedis}d that use of the incertae sedis
semantics results the taxonomy edge $AB$ grouping $A$ and $B$.  It also restores the
names $AB$ and $A$.  We may thus determine that $C$ has been placed within $A$ (and
its ancestor $AB$).

\subsection{Conflict with incertae sedis taxa}

When only tip nodes on the taxonomy are \emph{incertae sedis}, we have
a very simple form of the \emph{incertae sedis} supertree problem.
However, in practice entire clades may be \emph{incertae sedis}, and
so more complex issue arise.
In this section we consider a number of
cases that must be handled when attempting to place \emph{incertae
sedis} taxa.

\emph{Note that the synthesis of all input trees before the taxonomy
is unaffected by incertae sedis information.}

\subsection{Conflicting placement among input trees}

{[}\textbf{BDR:} \emph{maybe we should merge this with Figure.
\ref{fig:Broken-incertae-sedis}.
Although the point of figure
\ref{fig:placement-example1} is to talk about sub-problem
decomposition as well as what the result looks like.}{]}

The addition of \emph{incertae sedis} taxa allows new types of
conflict between input trees.
For example, different input trees might
place an incertae sedis taxon in conflicting locations.
This is
illustrated in Figure \ref{fig:placement-example1}, where the IS
taxon (ott7,ott6)ott10 is placed as sister to ott1 by phylogeny
$\tau_{1}$ and as sister ott3 by phylogeny $\tau_{2}$.

When this happens, the placement of the IS taxon is not influenced by
its being marked IS on the taxonomy.
Thus, in Example 1, the higher
ranked tree $\tau_{1}$ will be reflected in the synthesis tree, ott10
will be placed as sister to ott1.
In contrast, the conflicting
placement in $\tau_{2}$ will not be reflected in the synth tree.

All this would occur in the previous version of propinquity.
Where the
updated version differs that (a) the names ott9 and ott8 are retained
instead of being dropped.
(b) as a result of not breaking ott8 and
ott9, we do not move ott3 and ott2 up to ott11.
{[}BDR: Extend more figures!{]}

\subsection{Case 1: An \emph{incertae sedis} clade}

When a clade is marked as \emph{incertae sedis}, we need to consider
two cases.
In the first case, the clade may be placed within a sister
clade intact (Figure \ref{fig:Handling-incertae-sedis}).
However, if
the clade is not monophyletic in the synthesis tree, then we allow the
members of the clade to be placed separately (Figure \ref{fig:Broken-incertae-sedis}).
\begin{figure}
\hfill{}\subfloat[Input tree \#1]{\includegraphics[scale=0.5]{Figures/is-clade/tree1-shaded}

}\hfill{}\subfloat[Input tree \#2]{\includegraphics[scale=0.5]{Figures/is-clade/tree3-shaded}

}\hfill{}

\hfill{}\subfloat[Taxonomy]{\includegraphics[scale=0.5]{Figures/is-clade/tax-shaded}

}\hfill{}\subfloat[Synthesis Tree \#1]{\includegraphics[scale=0.5]{Figures/is-clade/synth2-shaded}

}\hfill{}

\caption{\label{fig:Broken-incertae-sedis} \emph{Incertae sedis} clade $C$
broken by conflicting placement.
Input trees (a) and (b) conflict
in the placement of taxa in $C$.
The synthesis supertree (d) places
places $C1$ and $C2$ separately within $A$, while unplaced taxa
float upwards.
(\textbf{Note:} This is what the sub-problem solver
does.
The entire pipeline would attach $C3$ at $A$ actually, since
it is the MRCA of $A1$ and $A2$.
Should we change this?) }

\end{figure}

When the input phylogenies conflict on where members of an incertae
sedis clade should be placed, neither placement is rejected.
Instead,
the taxon $C$ is broken, the name $C$ disappears, and $C3$ floats to
the top level.
Algorithmic, alternatively placements of children of an
incertae sedis clade lead to contesting of all edges along with taxa
might be placed, creating a large, merged sub-problem, as in Figure
\ref{fig:placement-example1}.

In Figure \ref{fig:Broken-incertae-sedis}, tree $T_{1}$ places $C1$
next to $A1$, while $T_{2}$ places $C2$ next to $A2$.
Since different
members of $C$ are placed, neither placement for $C$ is rejected.
Instead the taxon $C$ is broken, the name $C$ disappears, and $C3$
floats to the top level.
Furthermore, since taxon $C$ is a broken
\emph{incertae sedis }taxon, all of its children are effectively
incertae sedis independently, with the exception that they cannot be
placed within each other.
Therefore, the names $A$ and $AB$ are not
lost, since the taxa placed within them are \emph{incertae sedis}
names.

The situation here would be different if the monophyly of $C$ was
supported by an high-ranked input phylogeny.
In that case, instead of
breaking $C$, we would choose the placement in the highest-ranked
tree.

\subsection{Case 2: A nested \emph{incertae sedis} clade}

\begin{figure}
\subfloat[Input tree \#1]{\includegraphics[scale=0.5]{Figures/is-nested/tree3-shaded}

}\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/is-nested/tax-shaded}

}

\subfloat[Synthesis tree (\emph{incertae sedis} ignored)]{\includegraphics[scale=0.5]{Figures/is-nested/synth3-no-is-shaded}

}\hfill{}\subfloat[Synthesis tree (\emph{incertae sedis} considered)]{\includegraphics[scale=0.5]{Figures/is-nested/synth3-is-shaded}

}

\caption{\label{fig:Nested-incertae-sedis}Nested incertae sedis taxa.
(a)
Input tree \#1 places CI1 within CG, and C within $A$.
(b) In the
taxonomy, CI is \emph{incertae sedis }within C, while C is \emph{incertae
sedis} under the root.
(c) The synthesis tree when \emph{incertae
sedis} taxa are not considered.
The names AB, A, and CG are lost.
(d) The synthesis tree when \emph{incertae sedis} taxa are considered.
The names AB, A, and CG are regained, as well as the edges leading
to AB and CG.}

\end{figure}
When \emph{incertae sedis} clades are allowed, it is possible for an
\emph{incertae sedis} taxon to be nested within another \emph{incertae
sedis} taxon.
In such a case we must allow the higher-ranked taxon to
be placed within a sibling, while preserving the right of the lower
ranked taxon to be placed within a lower-level sibling (Figure
\ref{fig:Nested-incertae-sedis}).

\subsection{Case 3: Placement of one \emph{incertae sedis} taxon within
another.}
\subsection{Surprising results of the split-based semantics}

\subsubsection{Inter-digitation}
1. Inter-digitation

\bdrcomment{Ben: discuss figure \ref{fig:interdigitation}.}

\subsubsection{Placing one \emph{incertae sedis} group within another}
It would be nice if our semantics of incertae sedis taxa would allow us to place
one \emph{incertae sedis} sibling inside another.
A special case of
this is when a higher-rank incertae sedis taxon $A$ is placed as
sister to an incertae sedis taxon $B$, and then $B$ is placed within
$A$ (Figure \#).

\begin{figure}
\subfloat[Input tree \#1]{\includegraphics[scale=0.5]{Figures/is-within-is/tree1\lyxdot tre}

}\subfloat[Taxonomy]{\includegraphics[scale=0.5]{Figures/is-within-is/tax-shaded}

}\subfloat[Synthesis tree]{\includegraphics[scale=0.5]{Figures/is-within-is/synth1-shaded}

}

\caption{\label{fig:Placing-one-incertae}Placing one incertae sedis group
within another.
Here I is placed within A, and then AI is placed within
A.}
\end{figure}

2. Placing A next to B, then B within A.

\bdrcomment{Ben: Write discussion of Figure \ref{fig:Placing-one-incertae}}

\subsection{Splits-based semantics versus edit-based semantics}

Here we focus on a split-based semantics instead of a semantics based on applying
of a series of edit operations to the tree. The split-based semantics allows us
to seek trees that are simultaneously consistent with a set of split-constraints.
The use of a split-based semantics does impose some limitations. These come
primarily because in the edit-based semantics, after we place an
\emph{incertae sedis} taxon, it is no loner \emph{incertae sedis}, whereas with
the split-based semantics taxa are either always \emph{incertae sedis} or never
\emph{incertae sedis}.

This has an effect, for example, when we consider two incertae sedis siblings $A$
and $B$. If we can place $A$ within $B$, and $B$ within $A$, then a split- based
semantics allows us to freely inter-digitate $A$ and $B$. If we do not want to
allow $A$ and $B$ to be inter-digitated, then we cannot both $A$ within $B$ and $B$
within $A$. In addition, we can imagine that a family $A$ is incertae sedis within a
kingdom. In an edit-based semantics, we could place $A$ as a sibling to a genus $B$,
mark $A$ as non-\emph{incertae-sedis}, and then place $B$ within $A$. In a split-based
semantics, we must disallow either $B$ within $A$ or $A$ within $B$ if we want to disallow
interdigitation.

As a result of this, we cannot satisfy all five of these criteria. Thus, we
describe two alternative split-based semantics. The non-interdigitation
semantics gets \# (1,2,4,5), and the interdigitation semantics gets (1,3,4,5).
This makes the point that each semantics results from different choices about
which criteria to satisfy.

\section{Handling \emph{incertae sedis} taxa in the propinquity
pipeline}

In order to handle \emph{incertae sedis} taxa within propinquity, we
must modify some of the stages of the propinquity pipeline.
Subproblem
decomposition must place \emph{incertae sedis} taxa in the correct
subproblem.
Subproblem files must indicate which taxa are incertae
sedis.
The subproblem solver must read this information, account for
\emph{incertae sedis} taxa when solving subproblems, and correctly
name taxa that have been modified by having \emph{incertae sedis} taxa
place inside them.
The unpruner must be aware of \emph{incertae sedis}
taxa.
Annotations of the tree must be aware of \emph{incertae sedis}
taxa so that it does not consider taxa broken when they have an
incertae sedis taxon placed inside them.

\subsection{Exemplifying taxa}

One current problem is that well-known taxa like Fungi or Mammalia
tend to have a very large number of incertae sedis children, making
browsing in the tree viewer difficult.
This can happen when, for
example, fossils or other hard-to- place taxa get classified only to
the level of these well-known nodes and no further.
This leads to a
situation where well-known taxa serve as a dumping ground for unplaced
taxa.

Our current approach to this problem is to perform a second round of
pruning, or ``cleaning'', during the exemplification step.
Incertae
sedis taxa are pruned at this stage if they do not occur in any input
trees.
We thus generate a second ``cleaned taxonomy'' that has
undergone this further round of cleaning.
This approach improves on
the previous approach in that \emph{incertae sedis} taxa in input
trees are no longer pruned.
This approach also removes tons of
\emph{incertae sedis} children from nodes like ``Fungi'', where a lot
of unplaced fossils with few observable characters have been dumped.

However, this approach has the negative effect of pruning some
incertae sedis taxa that need not be pruned.
For example, suppose
\emph{incertae sedis} taxon $A$ contains 5 children, of which only 1
child $A_{1}$ occurs in an input tree.
If the taxon $A$ is not broken,
then it should be possible to attach the other 4 members of $A$ next
to $A_{1}$, without cluttering up the synthesis tree.
Such taxa have
been successfully placed even though they are not in any input tree.
This can only be discovered after synthesis is complete, though.

We should be able to to figure out which incertae sedis taxa are
placed (see section \ref{subsec:Placement}), and then mark the other
ones incertae sedis in the \emph{synthesis tree}.
We could then filter
out the taxa that \emph{remain }incertae sedis (or incertae\_sedis
inherited) in the tree viewer instead of the synthesis pipeline.

\subsection{Sub-problem decomposition}

The presence of \emph{incertae sedis }taxa poses a problem to sub-
problem decomposition, since taxonomy edges no longer completely
separate subproblems.
Instead, \emph{incertae sedis} taxa may attach
on either side of a taxonomy edge.
We seek to place \emph{incertae
sedis} taxa into subproblems in such a way that the subproblem solver
can perform the placement inside the subproblem.
This approach
postpones handling of conflict in \emph{incertae sedis} taxa to the
subproblem solver, where the problem is well formulated in terms of
splits.
However, it does have the effect of creating larger
subproblems.

We must also handle conflicting placements of \emph{incertae sedis}
taxa by different input trees.
Thus, if one input tree places the
\emph{incertae sedis} taxon $X$ in $((X)B)A$ and another places $X$ in
$((X)C)A$ then we must mark both edges $B$ and $C$ as contested edges,
even if these edges would \emph{not} be contested were taxon $X$ to be
removed.
This results in a new way to contest edges that involves the
interaction of two input trees, and not just the interaction of each
input tree with the taxonomy.

We choose to solve these problems by merging any subproblems that an
\emph{incertae sedis} taxon might be placed in.
The simplest way to
achieve this is simply to regard any taxon that has an \emph{incertae
sedis }taxon placed within it as contested.
This results in marking
both $B$ and $C$ as contested edges in the example above.
In fact,
this is the current behaviour of the non-\emph{incertae-sedis} aware
subproblem decomposer.
One downside of this approach is that, if we
have $((X)B)A)$ in one input tree, and $X$ is mentioned nowhere else,
then by marking $B$ as contested, we are creating unnecessarily large
subproblems.
We could instead placed $X$ in $B$ and avoid contesting
the edge $B$.
However, this approach is more complex and does not seem
necessary in practice.

\begin{figure}
\subfloat[Taxonomy $\protect\taxonomy$]{\includegraphics[scale=0.5]{Figures/placement1/tax\lyxdot tre}

}\subfloat[$\tau_{1}$]{\includegraphics[scale=0.5]{Figures/placement1/input1\lyxdot tre}

}\subfloat[$\tau_{2}$]{\includegraphics[scale=0.5]{Figures/placement1/input2\lyxdot tre}

}

\caption{\label{fig:placement-example1}Example.
An incertae sedis clade (ott6,ott7)
is placed in different subtrees by input trees $\tau_{1}$ and $\tau_{2}$.
In $\tau_{1}$, two nodes that correspond to the taxonomy their ingroup
extended to include (ott6,ott7), and the branches leading to these
nodes have been colored blue.
The dashed blue edge leads to a node
that is a newly-introduced degree-2 node which does not correspond
to any taxonomy node.
In $\tau_{2}$, only one node that corresponds
to a taxonomy node needs to have its ingroup extended.
The placement
of (ott6,ott7) into ott8 toward ott1 by $\tau_{1}$ conflicts with
the placement of (ott6,ott7) into ott9 toward ott3 by $\tau_{2},$}
\end{figure}
 For example, in Figure
\ref{fig:placement-example1}, input tree $\tau_{1}$ contests ott9 and
input tree $\tau_{2}$ contests ott9 and ott8.
Thus ott1, ott2, ott3,
ott6, and ott7 end up in the same sub-problem.

\subsection{Subproblem solution}

Our sub-problem solver naturally handles \emph{incertae sedis} taxa.
This is because we define the semantics of \emph{incertae sedis} taxa
in terms of partial splits, and our solver natively supports building
trees from partial splits through its use of the BUILD algorithm.
Handling \emph{incertae sedis} taxa thus requires loading incertae
sedis information and computing partial splits for \emph{incertae
sedis} taxa before solving a sub-problem.
After solving a sub-problem,
we must apply taxon names from the taxonomy tree to the sub-problem
solution tree.
The solution tree is considered to a fixed tree and not
to have any \emph{incertae sedis} nodes, or any other forms of
uncertainty.

\subsubsection{Reading incertae sedis information}

Currently, we read the \emph{incertae sedis} information as a list of
OTT ids for \emph{incertae sedis} taxa.
This does not require adding
further annotations to the node names.
Only taxonomy nodes can be
\emph{incertae sedis} at the moment, and only the taxonomy tree for
the subproblem contains OTT ids for internal nodes.
Therefore we
handle \emph{incertae sedis} information by constructing modified
split sets for the lowest-ranked tree when the list of \emph{incertae
sedis} nodes is not empty.

\subsubsection{Exclude sets modified by \emph{incertae sedis} marks}

Equation (\ref{eq:exsetformone}) leads to the following
algorithm to compute the exclude set for all nodes in a tree.
\begin{enumerate} \item Set the exclude set of the root node to be
empty \item For each \emph{node} (except the root) in preorder
\begin{itemize} \item combine the \emph{exclude} set of the parent
node with the \emph{include} set of non-\emph{incertae-sedis}
siblings.
\item store this set in a hash, with key \emph{node}
\end{itemize} \end{enumerate} This algorithm is currently implemented
in \emph{otc-solve- subproblem}.
We store the sets as \emph{std::set}.

\subsubsection{Implementation: finding the node for a name}

To find the node for a name $n$, we find the MRCA of the cluster
$S_{1}(n)$.
If the MRCA excludes the entire exclude group $S_{2}(n)$
then the name applies to the MRCA; otherwise the taxon does not exist
on the tree.

\subsubsection{Implementation: handling name clashes}

When multiple names $N=\{n_{1},\ldots n_{N}\}$ map to the same
solution node $x$, then these names must satisfy some tree structure
on the taxonomy, such that $n_{1}<n_{2}$ if $n_{1}$ is a descendant of
$n_{2}$ in the taxonomy.
If it is possible to find a name $n_{max}$
that is the unique maximal element of $N$, then it is permissible to
\begin{enumerate}
    \item create a monotypic parent $p(x)$ of $x$, and
assign $n_{max}$ to $p(x)$ 
    \item continue handling name clashes at $x$
with the set of possible names reduced to $N-n_{max}$.

\end{enumerate}
However, its certainly possible that there might not be any such
$N_{max}$, in which case we could just choose a name for $x$ from $N$
(perhaps not an \emph{incertae sedis} name) and then record all the
other names as equivalents somewhere.

\textbf{BDR:}\textbf{\emph{ }}\emph{we might get this behavior in a
nice an automatic way if we create a single fake leaf for each
monotypic taxonomy node that holds the node's leaf label.}

\subsubsection{Caveats}

When multiple I.S.
taxa have been moved to the root node of a
subproblem, they may be I.S.
over the entire subproblem, and some may
be I.S.
over others in an asymmetric manner.
Therefore, we might need
to specify additional information about the original attachment
location of the I.S.
taxa, such as their depth.
This only affects
problems that have been decomposed.

\textbf{BDR:} \emph{currently we don't actually move taxa to get them
into a subproblem.
So, is this even an issue?}

\subsection{Grafted supertree}

\emph{Question:} Does the synthesis tree contain any \emph{incertae
sedis} groups?\\ \emph{Answer:} The grafted supertree will not contain
any \emph{incertae sedis} groups.
However, when we attach pruned nodes
to a parent in the grafted supertree, we could mark such nodes
\emph{incertae sedis} if we want.

\subsection{Unpruning}

Currently the unpruner \emph{does not} require that the OTT ids are
named in the grafted solution before unpruning starts.
According to
Mark's document, he wasn't sure if such names were generated for nodes
that had an IS taxon placed inside of them, so otc-unprune-solution-
and-name-unnamed nodes throws away all the names and generates them
itself.

\textbf{BDR}: See document texttt{otcetera/doc/unprune-solution-and-name-unnamed-nodes.pdf}

The unpruner should record when unpruned nodes are \emph{incertae
sedis}.
Such nodes are unaffected by phylogenies, and so
\emph{incertae sedis} annotations for them make good sense.

\subsection{Annotation}

Annotation primarily involves running a conflict analysis between the
synthesis tree and each input tree.
Since neither tree has any
\emph{incertae sedis} taxa, the conflict algorithm does not need to
change.
Furthermore, if we allow \emph{incertae sedis} taxa that are
taxonomy-only to be annotated as \emph{incertae sedis} on the synth
tree, then such groups will not affect conflict with the input trees.
We would also like to allow running a conflict analysis between the
synthesis tree and the taxonomy tree.
However, naming the nodes
\emph{is} a (almost) run of conflict analysis on the taxonomy tree,
and this has already been done in a prior step.
So, the current
annotation procedure actually works as-is.

It would be nice to allow running conflict against the cleaned
taxonomy, though.
One way to do this would be to generated a ``placed
taxonomy'', with groups extended to include \emph{incertae sedis} taxa
that have been placed within them.
This would not require any updating
to the conflict-analysis code in the annotation step.
\begin{comment}

\subsection{Conflict service}

The current conflict service considers a group $A$ to conflict with
the taxonomy if group $A$ has an incertae sedis group $B$ placed
within it.
This doesn't affect the annotations, since taxon names are
added by the unpruner.
But it could make perfectly fine input trees
incorrectly look like they are the cause of broken taxa, if they
contain IS taxa.
Thus, it would be nice to have a modified conflict
algorithm.

\subsubsection{Current conflict algorithm}

The current conflict algorithm is pretty fast, but it works by
classifying tips into either (i) the include group or (ii) the exclude
group.
To avoid counting the exclude group for every split, we instead
count the total number of children for each node, and assume that any
children not in the include set are in the exclude set.
This is no
longer true when we have incertae sedis taxa.
I suspect that if we
want to handle incertae sedies, we'd need a third category (iii) for
``neither include group nor exclude group''.
\begin{enumerate}
 \item
Get induced trees on intersection of leaf sets \item Compute depth for
each node (nd->depth)
 \item Compute number of tips at or below each
node (nd->n\_tips)
 \item for each input tree node -> $nd$
 \begin{enumerate}
   \item skip the root \item skip monotypic \item if
its a tip then find corresponding (``terminal'') edges in synth tree
and continue \item leaves1 <- get the list of leaves in the include
group of $nd$ (in input) \item L2 <- find the total number of tips (L2
= sum {[}nd->n\_tips| nd <- leaves1{]})
   \item leaves2 <- get list of
corresponding synth leaf nodes (in synth)
   \item nodes <- find all
nodes between leaves2 and the MRCA (in synth)
   \item MRCA <- mrca of
leaves2 (in synth.
this uses the nd->depth annotation)
   \item Compute
number of tips in the include set (nd->include\_tips) below each node
in $nodes$ (in synth)
   \item if n\_include\_tips(MRCA) == n\_tips(MRCA)
then the MRCA displays $nd$
   \item if n\_include\_tips(MRCA) <
n\_tips(MRCA) then 
   \begin{itemize} 
     \item foreach node in nodes
      \begin{itemize}
         \item if (n\_include\_tips(nd) < n\_tips(nd) and
n\_include\_tips(nd) < l2)
          \begin{itemize}
            \item this is a conflict!
\end{itemize} 
           \end{itemize}
         \item if there are no conflicts, then this
is a resolved_by.

      \end{itemize}
\end{enumerate}
\end{enumerate}

\subsubsection{Modified conflict algorithm?}

This probably is outside the scope of the paper, but if we could come
up with a modified conflict algorithm, that would be nice/useful, and
probably novel.
It would also probably be slower...

\end{comment}

\section{Results}

Should we do this? We could say:  \begin{itemize} \item there are 331
incertae sedis taxa that are mentioned in input trees.
\begin{itemize}
\item what is the kingom / phylum / class / order / genus / species
\end{itemize} \item we placed 14 incertae sedis taxa inside a sister
taxon.
(``incertae sedis'', ``versus ``unplaced''?) \item we placed
$o$ incertae sedis taxa \emph{outside} all sister taxa.
\item we
confirmed $n$ incertae sedis taxa as being separate from all sister
(sampled) taxa?? \item we avoiding breaking 8 taxa that had IS taxa
placed inside them.
\item we allowed $z_{1}$ new taxa into the
synthesis tree that were incertae sedis.
\item we allowed $z_{2}$ new
taxa into the synthesis tree that are marked as extinct.
\item some
nodes have as many as $w$\emph{ incertae sedis} children, making them
unbrowseable when incertae sedis children are not excluded.
\item
$v_{1}$ input trees were previously excluded \emph{entirely} because
they are nested within in an incertae sedis taxon.
\item $v_{2}$ input
trees were previously excluded \emph{partially} because they are
nested within in an incertae sedis taxon.
\end{itemize} Currently the
numbers $z_{1}$ and $z_{2}$.

\section{Discussion}

One thing we could do (perhaps) that we are not currently doing, is to
have nodes marked as incertae sedis on the synth tree.
This would be
easy enough if such nodes are not affected in any way by the input
trees.
Thus, when unpruning nodes we could mark any nodes
\emph{incertae sedis} if they were marked \emph{incertae sedis} on the
taxonomy.

Secondly, I think we need to distinguish \emph{incertae sedis} taxa
that are ``unplaced'' from \emph{incertae sedis} taxa that do not
occur in any input tree.
I think that if $A$ contains child $A_{1}$
that is an input tree, and the taxon $A$ is not broken, then $A$ will
be placed, and thus any other children $A_{2},A_{3},\ldots,A_{n}$ will
also be placed, since they will be added as children of the (placed)
node $A$ by the unpruner.
This could be considered when deciding which
nodes to suppress in the tree viewer.

\paragraph{Taxonomy merging}

Not all \emph{incertae sedis} taxa in our taxonomy are directly
labeled \emph{incertae sedis} by taxonomists.
\emph{Incertae sedis}
taxa can also result from automatic merging of taxonomies to create
the OpenTree taxonomy.
For example, in Figure 3 of
\citet{rees2017automated}, cases \#4 and \#6 illustrate examples where
merging of two taxonomies leads to a taxonomy with a taxon of
uncertain placement.
The reason is primarily that if taxonomy
$\taxonomy_{1}$ contains more levels of hierarchy than taxonomy
$\taxonomy_{2}$, then we must add internal nodes to $\taxonomy_{2}$ to
align it to $\taxonomy_{1}$.
However, if taxonomy $\taxonomy_{2}$
contains more leaves than $\taxonomy_{1}$, then it is unclear if these
extra leaves should be nested inside the additional internal nodes, or
not.
Thus, the extra leaves are marked \emph{incertae sedis}.

For example, if $\taxonomy_{1}=((a,b)x,(c,d)y)z$ and
$\taxonomy_{2}=(a,b,c,d,e)z$ then $a$ and $b$ in $\taxonomy_{2}$
should be nested within $x$, but we do not know if $e$ should be
nested within $x$ or not.
Thus we obtain $((a,b)x,(c,d)y,?e)z$, where
$?$ indicates that taxon $e$ is marked \emph{incertae sedis}.

If incertae sedis taxa are cannot be placed within other incertae
sedis taxa, then such an approach does not work if $y$ is already
incertae sedis.
However, we take the approach that \emph{incertae
sedis} notations are not a general solution to expression of uncertain
placement.

\subsubsection{Comparison to operational definition}

The splits-based semantics for incertae sedis taxa has many desirable
properties.
In allows the use of the BUILD algorithm to assess
compatibility of split sets.
However, it also has some properties that
may not be expected.
For example, two incertae sedis siblings can be
freely interdigitated, which might not be the expected outcome.
The
current definition, also ignore ranks.
Thus if we had two incertae
sedis siblings that were genera, but other siblings were families,
then one might expect that the genus-level incertae sedis taxa could
be placed within families, but not intermixed.
Since the current
approach ignores ranks, it cannot do that.

The above definition of incertae sedis taxa does allow two incertae
sedis sister taxa to be interdigitated.
It also

Also, the splits-based semantics might not completely recover the
edit-operation semantics.
The corner cases are ...
.

\subsubsection{Intrusion into incertae sedis taxa}

In theory it could be possible for an \emph{incertae sedis} taxon $A$
to be placed within an \emph{incertae sedis} sibling $B$.
We initially
implemented this semantics.
It requires updating the recursion to
define $I(m,n)$ as the descendants of $m$ that can be access from $n$
without traversing an \emph{incertae sedis} node.

With a splits-based semantics, we cannot allow both (i) placing $A$ in
$B$ and (ii) placing $B$ in $A$ and also forbid interdigitating $A$
and $B$.

However, if we allow this behavior by reducing the exclude sets for
$A$ and $B$, then this has the unfortunate consequence that the
members of $A$ and $B$ may be inter-digitated.
In many cases,
\emph{incertae sedis} genera have siblings that are families, and the
expected semantics is that genera may be placed within their sibling
families, or may remain outside existing families.
In such cases,
inter-digitating the genera would not be expected behavior.
Thus, we
choose the semantics that \emph{incertae sedis} taxa exclude their
\emph{incertae sedis} siblings.
This is not because the other
semantics is impossible.
However, we note that use of the
interdigitating semantics allows a situation where the names for $A$
and $B$ both apply to the same node.

This semantics also complicates placement, since it is possible to
place $A$ as a sibling of $B$, and then place $B$ within $A$.
This
violates the current invariant that taxa are only ever placed within
more tipward taxa.
This makes the description of placement more
complicated, since we can no longer simply record all cases where a
parent node is not ancestor on the taxonomy.

\subsubsection{Usage in the literature}

The term \emph{incertae sedis} can be used in a number of different
ways.
For example,\emph{ }one author might describe \emph{incertae
sedis} genera that are asserted to be monophyletic, whereas another
author might describe \emph{incertae sedis} genera that may be
interdigitated.

\appendix
\section{Alternative semantics}
\subsubsection{semantics \#1b: modified to disallow inter-digitating
\emph{incertae sedis} siblings}

We additionally stipulate that it is not possible for \emph{incertae sedis} taxa
to intrude into their \emph{incertae sedis} siblings.  We thus define $\mathcal{I}(m,n)$ to
indicate descendants of $m$ that are excluded from $n$. Thus:
\begin{align*}
\mathcal{I}(m,n) & =
    \begin{cases}
        \emptyset & \text{if }n\text{ is \emph{incertae sedis}\,and }m\text{
is not}\\
        \mathcal{I}(m) & \text{otherwise}
    \end{cases}
\end{align*}

We can write the exclude set of a node $n$ in terms of
the exclude set of its parent:
\begin{eqnarray*}
    \mathcal{E}(n) & = \mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in
           siblings(n)\right].
           \label{eq:exsetformone}
\end{eqnarray*}
\subsubsection{An alternative semantics that allows placing an incertae sedis group within an incertae-sedis great-uncle}

\bdrcomment{This semantics is actually designed to allow the situation in Figure \ref{fig:Placing-one-incertae}.

  The problem is (might be) as follows: if we have the tree ``((?A1,A2)A,((D)C)B)MRCA;'', where A1 is incertae sedis, then we can place A1 within D.  This semantics does not exclude A1 from B, C, or D.  So, if A is broken, then nothing prohibits A1 from intruding into B, C, D, etc.

  The purpose of this was to allow placing B as a siblign to A1 and then placing A1 within B -- \emph{if} B were incertae sedis.  But B is not incertae sedis.

  We could try to fix up this situation by adding cases to the exclude rules to say that B excludes A1 if B is not incertae sedis.  But adding cases makes things messier.  I think it also does not actually solve the problem: suppose we mark B incertae sedis, but do NOT place it as sibling to A1.  Then again if A is broken, A1 can intrude into B.  We really want to allow A1 into B only if B is placed as sibling to A1, and I'm not sure how to do that with the partial-split-based semantics.  If you see a way, then great!
}

If we want to allow placing two incertae sedis siblings within each
other, then incertae sedis siblings are never excluded.

Furthermore, consider placing $A$ as a sibling of $B$, and then
placing $B$ within $A$.
In order for this to work, $A$ must not
exclude the incertae sedis children of $B$.

In order to accomplish both of these goals, we define the descendants
$\mathcal{I}^{\prime}(n)$ that are accessible without passing through
an incertae sedis node.

\begin{align*}
    \mathcal{I^{\prime}}(n) & =
    \begin{cases}
        \emptyset & \text{if }n\in\mathcal{U}\\
        \{n\} & \text{if }n\text{ is a leaf}\\
        [\mathcal{I}^{\prime}(c)|c\in children(n)] & \text{otherwise.}
\end{cases}
\end{align*} We may then define the taxa excluded from $n$
as:

\begin{eqnarray*} 
\mathcal{E}(n) & =\mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in siblings(n)\right].\label{eq:exsetformoneb}
\end{eqnarray*}

This semantics satisfies criteria 1,3,4, and 5 above.

\subsubsection{A third semantics}

\bdrcomment{I didn't really finish working this one out.}

OK, the idea here is that we want to avoid inter-digitation.
\emph{Incertae sedis} taxa thus exclude\emph{ incertae sedis}
siblings.
However, where the first semantics allows \emph{incertae
sedis }taxa to intrude into \emph{incertae sedis} descendants of
siblings, but not allow\emph{ incertae sedis} descendants of siblings
to intrude into \emph{incertae sedis} taxa, here we do the reverse.

OK, so the idea here is that an\emph{ incertae sedis} group excludes
\begin{itemize} \item \emph{incertae sedis} siblings \item
\emph{incertae sedis }descendants of ancestors \item but not\emph{
incertae sedis} descendants of non-incertae-sedis siblings
\end{itemize} It is also possible to allow placing $A$ as a sibling of
$B$ and then placing $B$ within $A$ if we disallow placing $A$ into
incertae sedis descendants of its siblings.
We may then allow placing
those descendants within $A$.

\begin{align*} \mathcal{I^{\prime}}(n) & =\begin{cases} \emptyset &
\text{if }n\in\mathcal{U}\\ \{n\} & \text{if }n\text{ is a leaf}\\{}
[\mathcal{I}^{\prime}(c)|c\in children(n)] & \text{otherwise.}
\end{cases} \end{align*}

We use $\mathcal{I}(m,n)$ to indicate the descendants of $m$ that are
excluded from $n$.
\begin{align*} \mathcal{I}(m,n) & =\begin{cases}
\emptyset & \text{if }n\text{ is \emph{incertae sedis}\,and }m\text{
is not}\\ \mathcal{I}(m) & \text{if }n\text{ and }m\text{ are both
}incertae\,sedis\\ \mathcal{I}^{\prime}(m) & \text{otherwise.}
\end{cases} \end{align*}

We may then define the taxa excluded from $n$ as:
\begin{align*}
\mathcal{E}(n) & =\begin{cases}
\left[\mathcal{L}-\mathcal{I}(parent(n))\right]\cup[\mathcal{I}(s,n)|s\in siblings(n)] & \text{if }n\in\mathcal{U}\\
\mathcal{E}(parent(n))\cup[\mathcal{I}(s,n)|s\in siblings(n)] & \text{otherwise.}
\end{cases}
\end{align*}
I think this works... but need to double-check.

Also: can we genericize $\mathcal{E}(parent(n))$?

\section{Details \& Methods}

\subsection{Cases of incertae sedis taxa}
1.1 Some examples in OTT This paragraph pulled from intro to
otcetera/doc /handling-incertae-sedis.pdf

The problem is becoming more acute because NCBI is putting a larger
number of taxa into groups that are marked as ``classified''. For
example, when OTT 2.9 was created NCBI's clasification of the bird
family Sylvidae included a group five genera that were placed within
\url{http://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi?mode=Undef&id=36270&lvl=3&lin=f&keep=1&srchmode=1&unlock||unclassified Sylviidae}.
This includes the genus Regulus. Thus in OTT 2.9 Regulus (ott ID = 3599326) is placed inside the family Silvidae (OTT ID =
259942).
The ``unclassified Sylviidae'' does not appear in OTT; instead
Regulus is flagged as ``unclassified,sibling higher'' and all of the
species within Regulus are flagged as ``unclassified inherited.''

The taxonomy (as of version 2.9) also contains 685 cases of taxa that
are flagged as both ``unclassified inherited'' and ``unclassified'' based
on {\tt grep unclass.*unclassified\_inherited taxonomy.tsv}.

We would like to stop suppressing (pruning) incertae sedis taxa, so
that groups such as Regulus can appear appear in the synthetic tree.

\bdrcomment{We get Regulus back, bu we do not end up placing it in Sylviidae
because Sylviidae is broken. It conflicts with about 20 different trees.}

\subsection{Taxonomy details}

\paragraph{Flags.}
The Open Tree's reference taxonomy is produced by a tool called
smasher that notices hints that a taxon is incertae sedis and labels
that taxon with one of five flags (see https://github.com/OpenTreeOfLife
/reference-taxonomy/wiki /Taxon-flags). This document is not concerned
with the flagging system per se, so ``incertae sedis'' will be used here
to refer to all of the flags that denote taxa with uncertain
placement.

\paragraph{Limited attachment range}
While a taxonomist may intend to use incertae sedis to
indicate a limited number of possible positions for a taxon to go in
the taxonomy, we do not retain any such details.

\paragraph{Container nodes.}
Some taxonomies have nodes with names like ``Incertae sedis (Bacteria)''.
Such nodes are called containers. They indicate that each child node is an
incertae sedis child of the container node's parent. The OTT retains container
nodes with the was\_container flag, but seems to have already moved all of the
container's children to the container's parent.

\section{Placements}
\textbf{Whale:}~ ?ott6145835~{[}6145835~Brandtocetus/genus{]}~<-~ott698424~{[}698424~Cetacea/orde r{]}~placed~under~ott4942423~{[}4942423~Cetotherium~/~genus{]}~<-~ott698424~{[}6 98424~Cetacea/order{]}

\textbf{Whale:}~?ott6145873~{[}6145873{]}~<-~ott698424~{[}698424~Cetacea~/~order{]}~placed~under~ott4942423~{[}4942423{]}~<-~ott698424~{[}698424{]}~

-~Cetacea~broken.

\textbf{Insect:}~?ott18090~{[}18090{]}~<-~ott672501~{[}672501{]}~placed~under~ot t275501~{[}275501{]}~<-~ott672501~{[}672501{]}

\textbf{Insect:}~?ott190269~{[}190269{]}~<-~ott672501~{[}672501{]}~placed~under~ ott275501~{[}275501{]}~<-~ott672501~{[}672501{]}~

\textbf{Insect:}~?ott349638~{[}349638{]}~<-~ott672501~{[}672501{]}~placed~under~ ott275501~{[}275501{]}~<-~ott672501~{[}672501{]}~

\textbf{Insect:}~?ott1009313~{[}1009313{]}~<-~ott672501~{[}672501~Reduviidae~/~f amily{]}~placed~under~ott275501~{[}275501{]}~<-~ott672501~{[}672501{]}~

-~Reduviidae~broken.

\textbf{Insect:~}?ott564110~{[}564110{]}~<-~ott939130~{[}939130~Thyrididae~/~family{]}~placed~under~ott757344~{[}757344{]}~<-~ott939130~{[}939130{]}~

\textbf{Plant:}~?ott679485~{[}679485{]}~<-~ott1074933~{[}1074933~Lamioidea~/~sub family{]}~placed~under~ott534796~{[}534796{]}~<-~

\textbf{Plant:}~?ott729184~{[}729184{]}~<-~ott1074933~{[}1074933~Lamioidea~/~sub family~{]}~placed~under~ott534796~{[}534796{]}~<-~ott1074933~{[}1074933{]}~

ott1074933~{[}1074933{]}~

-~Lamioidea~\emph{not}~broken

\textbf{Plant:}~?ott5144555~{[}5144555{]}~<-~ott23373~{[}23373{]}~placed~under~o tt447802~{[}447802{]}~<-~ott23373~{[}23373{]}

-~Orobanchaceae~broken

\textbf{Fly:}~?ott441413~{[}441413~Rhodesiella~lungliensis/species{]}~<-~ott1014 022~{[}1014022~Rubiaceae{]}~placed~under~ott571842~{[}571842~Pavetteae~/~tribe{ ]}~<-~ott997381~{[}997381~Ixoroidea/subfamily{]}~<-~ott1014022~{[}1014022~Rubia ceae/family{]}

\textbf{Plant:}~?ott31033~{[}31033{]}~<-~ott627035~{[}627035{]}~placed~under~ott 6019314~{[}6019314{]}~<-~ott627035~{[}627035{]}~

\textbf{Bacteria:}~?ott4018918~{[}4018918{]}~<-~ott5248084~{[}5248084{]}~placed~ under~ott167767~{[}167767{]}~<-~ott5248084~{[}5248084{]}~

\textbf{Bacteria:}?ott4792759~{[}4792759{]}~<-~ott5248084~{[}5248084{]}~placed~u nder~ott167767~{[}167767{]}~<-~ott5248084~{[}5248084{]}~

-~Erysipelotrichaceae~broken

\bibliographystyle{upmplainnat}
\bibliography{otcetera}

\end{document}